---
title: "Rmleda: Package Usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rmleda-package-usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

This vignette describes the use of the `supervised_data()`, `dftype()`, `autoimpute_na()`, and `dfscaling()` functions included in package `Rmleda`. The `Rmleda` package helps with preliminary EDA for a given dataset to perform various data preparation and wrangling tasks such as data splitting, exploration, imputation, and scaling. These functionalities were identified as commonly-performed tasks in supervised machine learning settings but may provide value in other project types as well.

### Installation

If you do not have the `devtools` package you can install it via cran

```{r devtools, eval=FALSE, echo=TRUE}
install.packages("devtools")
```

Then install Rmleda from GitHub as follows

```{r installation, echo=TRUE, message=FALSE, results='hide'}
devtools::install_github("UBC-MDS/Rmleda")
```
Lastly, load the package

```{r setup}
library(Rmleda)
```

### Dataset

To explore the functionality of Rmleda, we will use the following fictional dataset on chocolates!

```{r dataset}
toy_df <- data.frame(
  chocolate_brand = c('Lindt', 'Rakhat', 'Lindt', 'Richart', 'Not Available'),
  price = c(3, NA, 4, 6, 3),
  type = c("dark", "dark", "white", "white", "dark")
)1
```

### Data type exploration
The `dftype()` function will return the type of columns and variables for the input data frame. Furthermore, if there are non-numeric columns, it will return the unique values of non-numeric columns and their length.

```{r dftype}
dftype(toy_df)
```

The function returns a list of results. The first entity (`$summary`) is a summary of each column. the second entity(`$unique_df`) is a data frame that contains a non-numeric column name, unique values of each column, and its count.

We can access both summary and unique_df separately.
```{r dftype_result}
dftype(toy_df)$summary
dftype(toy_df)$unique_df
```

### Data splitting

After performing the above pre-processing tasks, we are ready to move on perform data splitting. Instead of manually creating variables to hold train_set, test_set, x_train, y_train, x_test and y_test, we can simply use the `supervised_data` function. We pass it `toy_df` along with the names of `x` features or columns and the name of the `y` target column as below. 

<!-- The function automatically calls [initial_split()]("https://rsample.tidymodels.org/reference/initial_split.html") from the `tidymodels` package with the default arguments and returns easy-access variables for train portion of the dataset (`train`), the test portion of the dataset (`test`), the train portion of the dataset containing `X` features only (`x_train`), the train portion of the dataset containing `y` targets only (`y_train`), the test portion of the dataset containing `X` features only (`x_test`), and the test portion of the dataset containing `y` targets only (`y_test`). You can also access the original dataset from the supervised_data object using the variable name `data`. -->

```{r function_usage}
super_data <-
  supervised_data(toy_df,
                  xcols = c('chocolate_brand', 'type'),
                  ycol = c('price'))
```

For example, we can quickly access the x_train portion

```{r train_subset}
super_data$train
```

and with just the x columns/feature

```{r x_train_subset}
super_data$x_train
```

Similarly we can access the other subsets of the data

```{r other_subsets, eval = FALSE}
super_data$data      # train portion of the dataset
super_data$train     # test portion of the dataset
super_data$test      # test portion of the dataset
super_data$x_train   # train portion of the dataset containing `x` targets only
super_data$y_train   # train portion of the dataset containing `y` targets only
super_data$x_test    # test portion of the dataset containing `x` targets only
super_data$y_test    # test portion of the dataset containing `y` targets only
```

Finally, we can customize the split by passing extra arguments to `supervised_data`. These will be passed to `initial_split()` from the `tidymodels` package. See the [documentation]("https://rsample.tidymodels.org/reference/initial_split.html") for more information.

```{r, additional_args}
super_data <-
  supervised_data(
    toy_df,
    xcols = c('chocolate_brand', 'type'),
    ycol = c('price'),
    prop = 1 / 2 # Change split ratio to 0.5
  ) 
```

And access the subsets in the same manner described previously

```{r custom_split_output}
# Note for datasets with sizes that result in non-integer splits, 
# the training portion size will be rounded up and the test portion size will
# be rounded down as per initial_model() implementation from tidymodels
super_data$train
super_data$test
```


